---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Libraries used
```{r}
library(zip)
library(ggplot2)
library(dplyr)
```

## Downloading the data
```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",destfile = "data.zip")
unzip("data.zip")
data<-read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
sumStep<-tapply(data$steps, data$date, sum)
```
#### Histogram of total number of steps taken each day
```{r histogram1}
hist(sumStep, main = "Total Steps per Day", xlab = "Steps")
```

#### Mean of the steps 
```{r}
mean(sumStep, na.rm = TRUE)
```

#### Median of the steps
```{r}
median(sumStep,na.rm = TRUE)
```

## What is the daily average activity pattern?

#### Time series plot
```{r timeseriesplot1}
stepInterval<-tapply(data$steps, data$interval, mean, na.rm=TRUE)

plot(as.numeric(names(stepInterval)),stepInterval,
     ylab = "Steps", 
     type = "l", 
     xlab = "Interval",
     main = "Average daily activity pattern")
```

#### Interval associated with maximum activity
```{r}
int<-names(stepInterval[which.max(stepInterval)])
step<-stepInterval[which.max(stepInterval)]

```
interval= `r int`
steps= `r step`

## Imputing the missing values

#### Calculating total number of missing values

```{r}
naval<-sum(is.na(data$steps))
print(naval)
```

#### New data set with missing values filled by mean of that interval

```{r}
splitdata<-split(data,data$interval)
for (i in 1:length(splitdata)){
  splitdata[[i]]$steps[is.na(splitdata[[i]]$steps)]<-stepInterval[i]
}
```

#### After filling of missing values, histogram and mean and median

```{r}
newData<-unsplit(splitdata,data$interval)
newSumStep<-tapply(newData$steps, newData$date, sum)
hist(newSumStep, main = "Total Steps per Day", xlab = "Steps")
m<-mean(newData$steps)
med<-median(newData$steps)
```

The new mean = `r m`
the new median= `r med`

## Are ther differences in activity patterns between weekdays na wwekends?

#### Creating two new factor variables "weekday" and weekend"
```{r}
newData$date<-as.POSIXct(newData$date)
newData$day<-""
newData[weekdays(newData$date)=="Saturday" | weekdays(newData$date)=="Sunday",4] <-"weekend"
newData[!(weekdays(newData$date)=="Saturday" | weekdays(newData$date)=="Sunday"),4]  <-"weekday"

```

#### Time series panel plot over weekdays and weekends
```{r weekday.endtimeseries}
weekdaymean<-with(newData[newData$day=="weekday",],tapply(steps, interval, mean, na.rm=TRUE))
weekendmean<-with(newData[newData$day=="weekend",],tapply(steps, interval, mean, na.rm=TRUE))

par(mfrow=c(1,2))

plot(as.numeric(names(weekdaymean)),weekdaymean,
     ylab = "Steps", 
     type = "l", 
     xlab = "Interval",
     main = "Average weekday activity pattern")

plot(as.numeric(names(weekendmean)),weekendmean,
     ylab = "Steps", 
     type = "l", 
     xlab = "Interval",
     main = "Average weekend activity pattern")

```










